### 2014-06-24 tex/main-body-text.tex
few changes before dinner

\subsection*{Analysis of coastal California steelhead} 
To compare the behavior of {\sc flock} to that of {\sc structure} we used a published 
dataset of 2596 steelhead \textit{Oncorhynchus mykiss}
genotyped at 15 microsatellite loci from \citet{Garzaetal_norcal}.  {\sc flock} 
was run six times using the default setting for all 
run parameters (initial partition = random, number of iterations = 20, number of runs = 50, 
LLOD threshold = 0) for values of $K$ from 2 to 6. The stopping point of 
$K=6$ was chosen after four consecutive values of $K$
yielded no plateau (see below). For each value of $K$ from 2 to 6, the no-admixture, 
uncorrelated-allele-frequency model implemented in {\sc structure} was run 6 times using 
50,000 sweeps of burn-in followed by 150,000 sweeps of sample collection from the posterior. 
All runs were performed on a JNCS D685 with an i7 3.40GHz processor and 8GB
of RAM running 64-bit Windows 7 OS. For each run {\sc flock} reports the 
normalized likelihood values (the exponenetiated log-likelihood value for an individuals 
membership to each $K$ clusters scaled to sum to one). 
This value is comparable to the $q_i$ values
returned by {\sc structure}, which, in the no-admixture case, are posterior probabilities of cluster membership.
The program {\sc clumpp} \citep{Jak&Ros2007} was used to relabel clusters among different runs,
while the program {\sc district} \citep{Rosenberg2004} was used to plot individual \textit{$q_i$} values.

\citet{Duc&Tur2012} promote a method for using {\sc flock} to estimate, $K$, the number of
subpopulations. The user must first specify the number of runs (different initial allocation of individuals 
to each cluster) as well as the number of iterations each individual has the opportunity to be reallocated 
to a different cluster. For each run, the log-likelihood value of membership to all $K$ clusters is computed
and the log likelihood difference (LLOD), the difference between the cluster with the highest likelihood 
and second highest likelihood, is reported for each individual. Runs that converge on the same solution 
are indicated by the same mean LLOD score and 
are referred to as a plateau. The plateau record is a list of how many runs ended up at these different solutions. 
\citet{Duc&Tur2012} developed two stopping rules that lead to the estimation of $K$. In the abscence of 
reasonable values for $k$ \citet{Duc&Tur2012} advocate running  {\sc flock} until either a single plateau 
$\geq 6$  is reached or four consecutive $K$ values yield no plateaus. If the first stopping condition is satisfied,
the $K$ for which there is a a single plateau $\geq 6$ is the point estimate for $K$. Alternatively, 
if multiple plateaus are observed with at least one plateau $\geq 6$, the largest value of $K$ that satisfies 
this conditon should be taken as the lower bound estimate of $K$. If no plateaus $\geq 6$ are observed, $K$
is \textit{undecided}.

@@

\subsection*{Results} 
Run times were relatively short for both programs. {\sc flock} 
required 1133 minutes to complete all the runs, averaging 188.8 minutes for 
each of six groups of runs over $K$ from 2 to 6. {\sc structure} completed all the runs in 
486 minutes with each of the six groups of runs averaging 80.8 minutes. Increasing
$K$ from of 2 to 6 increased {\sc flock}'s run time by a factor of 2.4, while the run time 
for  {\sc structure} increased by a factor of only 1.6.  
\hrule

The allocation of indivividuals into K clusters was congruent for each program. For all six groups 
of runs from \textit{k}=2-4, {\sc flock} lead to nearly identical clustering of individuals 
(See supplementary materials). It should be noted
that of the 50 starting allocations of individuals, the likelihood of membership for every individual 
to each cluster is only reported for the 'best run'. Of these 'best runs' for a \textit{k} of 2 only two  
ended with identical clustering (same mean LLOD score), all other runs 
differed by at most the allocation of 13 individuals. As \textit{k} increases so do the number of solutions
that {\sc flock} finds (Fig 1). These solutions, however, are suprisingly similar to
 the partions identified by {\sc structure}. Both programs idenfity clusters 
which can be attributed to the geographical boundaries identified by \citet{Garzaetal_norcal}.

>>>>

\subsection*{Analysis of coastal California steelhead} 
To compare the behavior of {\sc flock} to that of {\sc structure} we used a published 
dataset of 2596 steelhead \textit{Oncorhynchus mykiss}
genotyped at 15 microsatellite loci from \citet{Garzaetal_norcal}.  {\sc flock} 
was run six times using the default setting for all 
run parameters (initial partition = random, number of iterations = 20, number of runs = 50, 
LLOD threshold = 0) for values of $K$ from 2 to 6. The stopping point of 
$K=6$ was chosen after four consecutive values of $K$
yielded no plateau (see below). For each value of $K$ from 2 to 6, the no-admixture, 
uncorrelated-allele-frequency model implemented in {\sc structure} was run 6 times using 
50,000 sweeps of burn-in followed by 150,000 sweeps of sample collection from the posterior. 
All runs were performed on a JNCS D685 with an i7 3.40GHz processor and 8GB
of RAM running 64-bit Windows 7. For each set of runs {\sc flock} reports the 
normalized likelihood values (the exponenetiated log-likelihood value for an individuals 
membership to each $K$ clusters scaled to sum to one) for the run that belongs to 
the longest plateau - the 'best run'. 
This value is comparable to the $q_i$ values
returned by {\sc structure}, which, in the no-admixture case, are posterior probabilities of cluster membership.
The program {\sc clumpp} \citep{Jak&Ros2007} was used to relabel clusters among different runs,
while the program {\sc district} \citep{Rosenberg2004} was used to plot individual \textit{$q_i$} values.

\citet{Duc&Tur2012} promote a method for using {\sc flock} to estimate, $K$, the number of
subpopulations. The user must first specify the number of runs (different initial allocation of individuals 
to each cluster) as well as the number of iterations each individual has the opportunity to be reallocated 
to a different cluster. For each run, the log-likelihood value of membership to all $K$ clusters is computed
and the log likelihood difference (LLOD), the difference between the cluster with the highest likelihood 
and second highest likelihood, is reported for each individual. Runs that converge on the same solution 
are indicated by the same mean LLOD score and 
are referred to as a plateau. The plateau record is a list of how many runs ended up at these different solutions. 
\citet{Duc&Tur2012} developed two stopping rules that lead to the estimation of $K$. In the abscence of 
reasonable values for $k$, \citet{Duc&Tur2012} advocate running  {\sc flock} until either a single plateau 
$\geq 6$  is reached or four consecutive $K$ values yield no plateaus. If the first stopping condition is satisfied,
the $K$ for which there is a single plateau $\geq 6$ is the point estimate for $K$. Alternatively, 
if multiple plateaus are observed with at least one plateau $\geq 6$, the largest value of $K$ that satisfies 
this conditon should be taken as the lower bound estimate of $K$. If no plateaus $\geq 6$ are observed, $K$
is \textit{undecided}.

@@

\subsection*{Results} 
Run times were relatively short for both programs. {\sc flock} 
required 1133 minutes to complete all the runs, averaging 188.8 minutes for 
each of six groups of runs over $K$ from 2 to 6. {\sc structure} completed all the runs in 
486 minutes with each of the six groups of runs averaging 80.8 minutes. Increasing
$K$ from of 2 to 6 increased the run time for {\sc flock} by a factor of 2.4, while the run time 
for {\sc structure} increased by a factor of only 1.6.
\hrule

The allocation of indivividuals into K clusters was similar for each program. For all six groups 
of runs from \textit{k}=2-4, {\sc flock} lead to nearly identical clustering of individuals 
(See supplementary materials) none of which differed substantially from results obtain from 
{\sc structure} (Fig 1). Of the 'best runs' for the six groups for \textit{k} of 2, two  
ended with identical clustering (same mean LLOD score), all other runs 
differed by at most the allocation of 13 individuals. As 
\textit{k} increases so do the number of solutions
that {\sc flock} finds (Fig 1). For all froups of runs for $K$ $\geq 2$, no plateaus were observed - all intial 
allocations resulted in solutions that differed slightly. These solutions, however, are suprisingly similar to
 the partions identified by {\sc structure}. Both programs idenfity clusters 
which can be attributed to the geographical boundaries identified by \citet{Garzaetal_norcal}.
